---
title: "Tools to create summary tables and plots"
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    highlight: null
    number_sections: yes
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Tools to create summary tables and plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Note, this example is illustrated using fake data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r warning=FALSE, message=FALSE}
library(youthvars)
```

# Summarising descriptive data with standardised tables and plots
The youthvars package includes a number of convenience functions for creating descriptive statistics tables and plots. The tables and plots can be outputted in HTML, PDF / LaTeX and Word formats. Note that tables outputted to Word format require some additional manual editing (e.g. modifying table headers).

```{r}
output_type_1L_chr <- ifelse(knitr::is_html_output(),
                                          "HTML",
                                          "PDF")
```

## Input dataset
We are going to illustrate the summary table and plotting functions using the same [fake dataset that we used in other vignettes](Replication_DS.html).

```{r message=FALSE, results='hide', warning=FALSE}
data("replication_popl_tb")
data_tb <- replication_popl_tb %>%
  add_adol6d_scores(prefix_1L_chr = "aqol6d_q",
                    id_var_nm_1L_chr = "fkClientID",
                    wtd_aqol_var_nm_1L_chr = "aqol6d_total_w")
```

Some of the functions we are illustrating in this vignette also require that the dataset has a dictonary of the `ready4use_dictionary` class.

```{r }
data("repln_ds_dict_r3", package = "youthvars")
data("aqol_scrg_dict_r3", package = "youthvars")
dictionary_tb <- ready4::renew(repln_ds_dict_r3,
                                      new_cases_r3 = aqol_scrg_dict_r3)
```

Currently, youthvars plotting and tabular functions are designed for use with datasets that have the following structural properties:

 - include a unique client identifier variable;
 - include data from two time-points, with data collection round denoted by a  factor variable two levels; and
 - is in long format (ie where measures at different time-points for the same individual are stacked on top of each other in separate rows). 

To check whether these conditions hold for your dataset, you can use the `assert_ds_is_valid` function. The function will return an informative error message if the conditions do not hold.

```{r}
assert_ds_is_valid(data_tb,
                   id_var_nm_1L_chr = "fkClientID",
                   round_var_nm_1L_chr = "round",
                   round_bl_val_1L_chr = "Baseline")
```

Our dataset passed the above test but if we set the unique identified variable to the same value for each record, then an error message is triggered.

```{r error=TRUE}
assert_ds_is_valid(data_tb %>% dplyr::mutate(fkClientID=1),
                   id_var_nm_1L_chr = "fkClientID",
                   round_var_nm_1L_chr = "round",
                   round_bl_val_1L_chr = "Baseline")
```

Similarly, if our dataset contained three rounds of data collection rather than two, it would also fail.

```{r error=TRUE}
assert_ds_is_valid(data_tb %>% dplyr::mutate(round=sample(c("Baseline","Mid","End"), size = nrow(data_tb), replace = T)),
                   id_var_nm_1L_chr = "fkClientID",
                   round_var_nm_1L_chr = "round",
                   round_bl_val_1L_chr = "Baseline")
```

## Descriptive statistics tables
We can use a combination of the `make_descv_stats_tbl` and `print_descv_stats_tbl` functions to summarise specified variables of interest. The `make_descv_stats_tbl` function makes a call to the `tableby` function from the `aresenal` package and adapts the output of that call.

```{r participantstb, tab.cap='Participant characteristics by data collection round', tab.id = 'participantstb', results="asis"}
variable_nms_chr <- c("d_age","d_sexual_ori_s",
                      "d_ATSI","d_studying_working","d_relation_s")
make_descv_stats_tbl(data_tb = data_tb, 
                     key_var_nm_1L_chr = "round",
                     dictionary_tb = dictionary_tb,
                     variable_nms_chr = variable_nms_chr,
                     nbr_of_digits_1L_int = 2) %>%
  print_descv_stats_tbl(bl_fup_vals_chr = c("Baseline","Follow-up"),
                        output_type_1L_chr = output_type_1L_chr,
                        round_var_nm_1L_chr = "round",
                        variable_nms_chr = variable_nms_chr)
```

Minor modifications to the arguments supplied to these function can provide insight as to wheter variable distribution parameters differed by data collection round.

```{r outcomestb, tab.cap='Outcomes by data collection round', tab.id = 'outcomestb', results="asis"}
variable_nms_chr <- c("k6_total", "phq9_total", "bads_total", "gad7_total")
make_descv_stats_tbl(data_tb = data_tb, 
                     key_var_nm_1L_chr = "round",
                     dictionary_tb = dictionary_tb,
                     variable_nms_chr = variable_nms_chr,
                     test_1L_lgl = T,
                     nbr_of_digits_1L_int = 2) %>%
  print_descv_stats_tbl(bl_fup_vals_chr = c("Baseline","Follow-up"),
                        output_type_1L_chr = output_type_1L_chr,
                        round_var_nm_1L_chr = "round",
                        test_1L_lgl = T,
                        variable_nms_chr = variable_nms_chr)
```

Further minor changes to the values supplied to function arguments and an additional call to the `add_participation_var` function, can provide insight as to whether participants supplying data at both timepoints differed from those who supplied only baseline data.

```{r participtb, tab.cap='Outcomes by data completeness', tab.id = 'participtb', results="asis"}
data_tb <- data_tb %>%
  add_participation_var() 
make_descv_stats_tbl(data_tb = data_tb  %>%
                             dplyr::filter(round == "Baseline"), 
                           key_var_nm_1L_chr = "participation",
                           dictionary_tb = dictionary_tb,
                           variable_nms_chr = variable_nms_chr,
                           test_1L_lgl = T,
                           nbr_of_digits_1L_int = 2) %>%
  print_descv_stats_tbl(bl_fup_vals_chr = c("Baseline","Follow-up"),
                        header_col_nms_chr = data_tb$participation %>% unique(),
                        output_type_1L_chr = output_type_1L_chr,
                        round_var_nm_1L_chr = "round",
                        test_1L_lgl = T,
                        variable_nms_chr = variable_nms_chr)
```

## Correlation tables
We can use a combination of the `transform_ds_for_tstng` and `make_corstars_tbl_xx` functions to summarise correlations of interest. The latter function adapts output from a call to the `rcorr` function from the `Hmisc` package.

```{r blcorstb, tab.cap='Correlations table (follow-up)', tab.id = 'blcorstb', results="asis"}
transform_ds_for_tstng(data_tb,
                       depnt_var_nm_1L_chr = "aqol6d_total_w",
                       depnt_var_max_val_1L_dbl = Inf,
                       candidate_predrs_chr = variable_nms_chr,
                       round_var_nm_1L_chr = "round",
                       round_val_1L_chr = "Baseline") %>%
  make_corstars_tbl_xx(result_chr = output_type_1L_chr)
```

We can repeat the preceding step for data from the follow-up collection round.

```{r fupcorstb, tab.cap='Correlations table (follow-up)', tab.id = 'fupcorstb', results="asis"}
transform_ds_for_tstng(data_tb,
                       depnt_var_nm_1L_chr = "aqol6d_total_w",
                       depnt_var_max_val_1L_dbl = Inf,
                       candidate_predrs_chr = variable_nms_chr,
                       round_var_nm_1L_chr = "round",
                       round_val_1L_chr = "Follow-up") %>%
  make_corstars_tbl_xx(result_chr = output_type_1L_chr)
```

# Plots
We can use the `make_itm_resp_plts` function to plot responses to all individual questions from a scale.

```{r itemsfig, fig.cap="AQoL-6D item responses", out.width = "600px", results='hide', message=FALSE}
plt <- make_itm_resp_plts(data_tb,
                   col_nms_chr = names(dplyr::select(data_tb,
                                                     starts_with("aqol6d_q"))),
                   lbl_nms_chr = c("Household tasks", "Getting around",
                                   "Morbility","Self care","Enjoy close rel\'s",
                                   "Family rel\'s", "Community involv\'t",
                                   "Despair","Worry", "Sad", "Agitated",
                                   "Energy level", "Control", "Coping",
                                   "Frequency of pain", "Degree of pain",
                                   "Pain interference","Vision", "Hearing",
                                   "Communication"),
                   plot_rows_cols_pair_int = c(5L,4L),
                   heights_int = c(20L, 1L),
                   round_var_nm_1L_chr = "round",
                   y_label_1L_chr = "")
```

Similarly, we can use the `make_sub_tot_plts` function to weighted sub-scale scores.

```{r domainsfig, fig.cap="AQoL-6D weighted domain scores", out.width = "600px",  results='hide', message=FALSE}
plt <- make_sub_tot_plts(data_tb,
                  col_nms_chr = names(dplyr::select(data_tb,
                                                    dplyr::starts_with("vD"))),
                  plot_rows_cols_pair_int = c(3L,2L),
                  round_var_nm_1L_chr = "round",
                  heights_int = c(10L, 1L),
                  y_label_1L_chr = "")
```

Finally, the `make_var_by_round_plt` function allows us to compare the distribution of a variable of interest by data collection round.

```{r totalsfig, fig.cap="AQoL-6D total scores by data collection round", out.width = "600px"}
make_var_by_round_plt(data_tb,
                      var_nm_1L_chr = "aqol6d_total_w",
                      round_var_nm_1L_chr = "round",
                      x_label_1L_chr = "AQoL-6D total weighted score")
```

