---
title: "Replication Dataset"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    includes:
      in_header: preamble.tex
    number_sections: no
vignette: >
  %\VignetteIndexEntry{Replication Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r message=F}
library(youthvars)
```
## Origin and purpose of the youthvars replication dataset
The replication dataset distributed with the youthvars package is a purely synthetic (entirely fake) dataset that is nonetheless representative (in terms of constituent variables and their uni-variate and [to some extent] joint distributions) of the original study dataset from which we developped the initial set of classes and functions included in youthvars. The purpose of the replication dataset is two-fold:

1) to help users replicate the results in our original study (by applying the same algorithms we used to a different but similar dataset, users will be able to get broadly similar results);

2) to help users trial the functions in the youthvars package (and the related TTU and youthu packages) before potentially applying these functions to their own data.

## Replication dataset dictionary

An overview of the contents of the replication dataset is provided in a data dictionary that can be accessed with the following command.

```{r message=FALSE}
data("repln_ds_dict_r3", package = "youthvars")
dictionary_tb <- repln_ds_dict_r3
```

\blandscape

```{r ddic, echo = knitr::is_html_output(), results='asis'}
dictionary_tb %>%
  ready4show::print_table(output_type_1L_chr = ifelse(knitr::is_html_output(),"HTML","PDF"), 
                          caption_1L_chr = "Data dictionary",  
                          use_lbls_as_col_nms_1L_lgl = T, 
                          hline_after_ls = c(-1),
                          add_to_row_ls = list(pos = list(-1,nrow(dictionary_tb)),
                                               command = c(Hmisc::label(dictionary_tb) %>%
                                                             Hmisc::latexTranslate() %>%
                                                             paste0(collapse = " & ") %>%
                                                             paste0("\\\\\n"),
                                                           paste("\\\\\n"))),
                          mkdn_tbl_ref_1L_chr = "tab:ddic")
```

\elandscape

## Accessing the replication dataset
The replication dataset itself can be accessed using the following command.

```{r}
data("replication_popl_tb", package = "youthvars")
raw_data_tb <- replication_popl_tb 
```

Adding variable labels makes it easier to interpret the dataset. In the following command, we add lables from the data dictionary to the replication dataset.

```{r }
raw_data_tb <- raw_data_tb %>%
  ready4use::add_labels_from_dictionary(dictionary_tb)
```

The first six cases from the replication dataset are reproduced below, using variable labels as column headings.

\blandscape

```{r repds, echo = knitr::is_html_output(), results='asis'}
raw_data_tb %>%
  head() %>%
  dplyr::mutate(d_interview_date = as.character(d_interview_date)) %>%
  ready4show::print_table(output_type_1L_chr = ifelse(knitr::is_html_output(),"HTML","PDF"),
                          caption_1L_chr = "Replication dataset", 
                          use_lbls_as_col_nms_1L_lgl = T, 
                          hline_after_ls = c(-1),
                          add_to_row_ls = list(pos = list(-1,nrow(raw_data_tb %>%
  head())),
                                               command = c(Hmisc::label(raw_data_tb) %>%
                                                             Hmisc::latexTranslate() %>%
                                                             paste0(collapse = " & ") %>%
                                                             paste0("\\\\\n"),
                                                           paste("\\\\\n"))),
                          mkdn_tbl_ref_1L_chr = "tab:repds") 
```
\elandscape
