---
title: "Quality Assure Youth Mental Health Dataset Variables"
output: rmarkdown::html_vignette
# output: 
#   rmarkdown::html_vignette:
#     toc: true
#   pdf_document:
#     highlight: null
#     number_sections: yes
vignette: >
  %\VignetteIndexEntry{Quality Assure Youth Mental Health Dataset Variables}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE, warning=FALSE}
library(ready4)
library(youthvars)
```

## Variable classes and data integrity
The `youthvars` package defines a number of [ready4 framework sub-module classes](https://www.ready4-dev.com/docs/framework/implementation/modularity/#ready4-model-sub-modules) that can be used to quality assure input or output datasets for the [ready4 economic model of youth mental health](https://www.ready4-dev.com/). `youthvars` sub-modules are potentially useful for:

1) facilitating automated data integrity checks that verify no impermissible values (e.g. utility scores greater than one) are present in source data, transformed data or results; and
2) automating the selection of the appropriate method to apply to each data type.

## Included sub-module classes

The initial set of sub-module classes included in the `youthvars` package are one class for Assessment of Quality of Life (Adolescent) health utility and one for each of the predictors used in the utility prediction algorithms included in the related [youthu](https://ready4-dev.github.io/youthu/) package. 

### Assessment of Quality of Life Six Dimension (Adolescent) Health Utility 

The `youthvars_aqol6d_adol`{.R} class is defined for numeric vectors with a minimum value of 0.03 and maximum value of 1.0.

```{r}
youthvars_aqol6d_adol(0.4)
```

```{r}
youthvars_aqol6d_adol(c(0.03,0.2,1))
```

Non numeric objects and values outside these ranges will produce errors.

```{r error = T}
youthvars_aqol6d_adol("0.5")
```

```{r error = T}
youthvars_aqol6d_adol(-0.1)
```

```{r, error = T}
youthvars_aqol6d_adol(1.2)
```

### Child Health Utility Nine Dimension - Australian Adolescent Scoring

The `youthvars_chu9d_adolaus`{.R} class is defined for numeric vectors with a minimum value of -0.2118 and maximum value of 1.0.

```{r}
youthvars_chu9d_adolaus(0.4)
```

```{r}
youthvars_chu9d_adolaus(c(0.03,0.2,1))
```

Non numeric objects and values outside these ranges will produce errors.

```{r error = T}
youthvars_chu9d_adolaus("0.5")
```

```{r error = T}
youthvars_chu9d_adolaus(-0.3)
```

```{r, error = T}
youthvars_chu9d_adolaus(1.2)
```

### Behavioural Activation for Depression Scale (BADS)

The `youthvars_bads`{.R} class is defined for integer vectors with a minimum value of 0 and maximum value of 150.

```{r}
youthvars_bads(143L)
```

```{r}
youthvars_bads(as.integer(c(1,15,150)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_bads(22.5)
```

```{r, error = T}
youthvars_bads(-1L)
```

```{r error = T}
youthvars_bads(160L)
```

### Generalised Anxiety Disorder Scale (GAD-7)

The `youthvars_gad7`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 21.

```{r}
youthvars_gad7(15L)
```

```{r}
youthvars_gad7(as.integer(c(0,14,21)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_gad7(14.6)
```

```{r error = T}
youthvars_gad7(-1L)
```

```{r error = T}
youthvars_gad7(22L)
```

### Kessler Psychological Distress Scale (K6) - Australian Scoring System

The `youthvars_k6_aus`{.R} class is defined for integer vectors with a minimum value of 6 and a maximum value of 30.

```{r}
youthvars_k6_aus(21L)
```

```{r}
youthvars_k6_aus(as.integer(c(6,13,25)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_k6_aus(11.2)
```

```{r error = T}
youthvars_k6_aus(1L)
```

```{r error = T}
youthvars_k6_aus(31L)
```

### Kessler Psychological Distress Scale (K6) - US Scoring System

The `youthvars_k6`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 24.

```{r}
youthvars_k6(21L)
```

```{r}
youthvars_k6(as.integer(c(0,13,24)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_k6(11.2)
```

```{r error = T}
youthvars_k6(-1L)
```

```{r error = T}
youthvars_k6(25L)
```

### Kessler Psychological Distress Scale (K10) - Australian Scoring System

The `youthvars_k10_aus`{.R} class is defined for integer vectors with a minimum value of 10 and a maximum value of 50.

```{r}
youthvars_k10_aus(21L)
```

```{r}
youthvars_k10_aus(as.integer(c(13,25,41)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_k10_aus(11.2)
```

```{r error = T}
youthvars_k10_aus(9L)
```

```{r error = T}
youthvars_k10_aus(51L)
```

### Kessler Psychological Distress Scale (K10) - US Scoring System

The `youthvars_k10`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 40.

```{r}
youthvars_k10(21L)
```

```{r}
youthvars_k10(as.integer(c(0,13,34)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_k10(11.2)
```

```{r error = T}
youthvars_k10(-1L)
```

```{r error = T}
youthvars_k10(41L)
```

### Overall Anxiety Severity and Impairment Scale (OASIS)

The `youthvars_oasis`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 20.

```{r}
youthvars_oasis(15L)
```

```{r}
youthvars_oasis(as.integer(c(0,12,20)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_oasis(14.2)
```

```{r error = T}
youthvars_oasis(-1L)
```

```{r error = T}
youthvars_oasis(21L)
```
### Patient Health Questionnaire (PHQ-9)

The `youthvars_phq9`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 27.

```{r}
youthvars_phq9(11L)
```

```{r}
youthvars_phq9(as.integer(c(0,13,27)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_phq9(15.2)
```
```{r, error = T}
youthvars_phq9(-1L)
```
```{r error = T}
youthvars_phq9(28L)
```

###  Screen for Child Anxiety Related Disorders (SCARED)

The `youthvars_scared`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 82.

```{r}
youthvars_scared(77L)
```

```{r}
youthvars_scared(as.integer(c(0,42,82)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_scared(33.2)
```
```{r, error = T}
youthvars_scared(-1L)
```
```{r error = T}
youthvars_scared(83)
```
###  Social and Occupational Functioning Assessment Scale (SOFAS)

The `youthvars_sofas`{.R} class is defined for integer vectors with a minimum value of 0 and a maximum value of 100.

```{r}
youthvars_sofas(44L)
```

```{r}
youthvars_sofas(as.integer(c(0,23,89)))
```

Non-integers and values outside these ranges will produce errors.

```{r, error = T}
youthvars_sofas(73.2)
```
```{r, error = T}
youthvars_sofas(-1L)
```
```{r error = T}
youthvars_sofas(103L)
```
