---
title: "Describe Datasets With YouthvarsProfile and YouthvarsSeries"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Describe Datasets With YouthvarsProfile and YouthvarsSeries}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

Note: **This vignette is illustrated with fake data**. The dataset explored in this example should not be used to inform decision-making.

```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(youthvars)
```

Youthvars provides a two classes - `YouthvarsProfile` and `YouthvarsSeries` that are useful for describing features of datasets. The tools in `youthvars` build on the metadata included in a [Ready4useDyad](https://ready4-dev.github.io/ready4use/articles/V_02.html).

## Ingest data
To start we ingest  `x`, a `Ready4useDyad` (dataset and data dictionary pair) that we can download from a remote repository.

```{r}
x <- ready4use::Ready4useRepos(dv_nm_1L_chr = "fakes",
                               dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/W95KED",
                               dv_server_1L_chr = "dataverse.harvard.edu") %>%
  ingest(fls_to_ingest_chr = "ymh_clinical_dyad_r4",
         metadata_1L_lgl = F)
```

## Add metadata
We could add metadata about `x`, such as the unique identifier variable name, by transforming it to a `YouthvarsProfile` instance.

```{r}
## Not run
# x <- YouthvarsProfile(a_Ready4useDyad = x,
#                       id_var_nm_1L_chr = "fkClientID")
```

However, in this case the data we ingested includes a longitudinal dataset. It is therefore preferable to transform `x` into a `YouthvarsSeries` instance.  `YouthvarsSeries` objects contain all of the fields of `YouthvarsProfile` objects, but also include additional fields that are specific for longitudinal datasets (e.g. `timepoint_var_nm_1L_chr` and `timepoint_vals_chr` that respectively specify the data-collection timepoint variable name and values and `participation_var_1L_chr` that specifies the desired name of a yet to be created variable that will summarise the data-collection timepoints for which each unit record supplied data).

```{r}
x <- YouthvarsSeries(a_Ready4useDyad = x,
                     id_var_nm_1L_chr = "fkClientID",
                     participation_var_1L_chr = "participation",
                     timepoint_vals_chr = c("Baseline","Follow-up"),
                     timepoint_var_nm_1L_chr = "round")
```

## YouthvarsSeries methods
Currently, only methods for `YouthvarsSeries` (and not yet `YouthvarsProfile`) have been included with the `youthvars` package. These methods are summarised in the following sections.

### Validate data
We use the `ratify` method to ensure that `x` has been appropriately configured for methods examining datasets reporting measures at two timepoints.

```{r}
x <- ratify(x,
            type_1L_chr = "two_timepoints")
```

### Inspect data
We can now specify the variables that we would like to prepare descriptive statistics for using the `renewSlot` and `renew` methods. The variables to be profiled are specified in arguments beginning with "compare_". Use `compare_ptcpn_chr` to compare variables based on whether cases reported data at one or both timepoints and `compare_by_time_chr` to compare the summary statistics of variables by timepoints, e.g at baseline and follow-up. If you wish these comparisons to report p values, then use the `compare_ptcpn_with_test_chr` and `compare_by_time_with_test_chr` arguments.

```{r}
x <- renewSlot(x,
               compare_by_time_chr = c("d_age","d_sexual_ori_s","d_studying_working"),
               compare_by_time_with_test_chr = c("k6_total", "phq9_total", "bads_total"),
               compare_ptcpn_with_test_chr = c("k6_total", "phq9_total", "bads_total")) %>%
  renew(type_1L_chr = "characterize")

```

The tables generated in the preceding step can be inspected using the `exhibit` method.

```{r participtb, tab.cap='Outcomes by data completeness', tab.id = 'participtb', results="asis"}
x %>%
  exhibit(type_1L_chr = "characterize",
          profile_idx_int = 1L,
          output_type_1L_chr = "HTML")
```

```{r outcomestb, tab.cap='Outcomes by data collection round', tab.id = 'outcomestb', results="asis"}
x %>%
  exhibit(type_1L_chr = "characterize",
          profile_idx_int = 2L,
          output_type_1L_chr = "HTML")
```

```{r outcomestbtst, tab.cap='Outcomes by data collection round (with p values)', tab.id = 'outcomestbtst', results="asis"}
x %>%
  exhibit(type_1L_chr = "characterize",
          profile_idx_int = 3L,
          output_type_1L_chr = "HTML")
```

The `depict` method can create plots, comparing numeric variables by timepoint.

```{r totalsfig, fig.cap="SOFAS total scores by data collection round", out.width = "600px"}
depict(x,
       type_1L_chr = "by_time",
       var_nms_chr = c("c_sofas"),
       label_fill_1L_chr = "Time",#
       labels_chr = c("SOFAS"),#
       y_label_1L_chr = "")
```
