---
title: "Describe And Disseminate Youth Mental Health Datasets"
output: rmarkdown::html_vignette
# output: 
#   rmarkdown::html_vignette:
#     toc: true
#   pdf_document:
#     highlight: null
#     number_sections: yes
vignette: >
  %\VignetteIndexEntry{Describe And Disseminate Youth Mental Health Datasets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

Note: **This vignette is illustrated with fake data**. The dataset explored in this example should not be used to inform decision-making.

```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE, warning=FALSE}
library(ready4)
library(youthvars)
```

Youthvars provides two [ready4 framework modules](https://www.ready4-dev.com/docs/framework/implementation/modularity/#ready4-model-modules) - `YouthvarsProfile` and `YouthvarsSeries` that form part of the [readyforwhatsnext economic model of youth mental health](https://readyforwhatsnext.org/). The ready4 modules in `youthvars` extend the [Ready4useDyad module](https://www.ready4-dev.com/docs/framework/use/authoring-data/label-data/) and can be used to help describe key structural properties of youth mental health datasets.

## Ingest data
To start we ingest  `X`, a `Ready4useDyad` (dataset and data dictionary pair) that we can download from a remote repository.

```{r}
X <- ready4use::Ready4useRepos(dv_nm_1L_chr = "fakes",
                               dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/W95KED",
                               dv_server_1L_chr = "dataverse.harvard.edu") %>%
  ingest(fls_to_ingest_chr = "ymh_clinical_dyad_r4",
         metadata_1L_lgl = F)
```

## Add metadata
If a dataset is cross-sectional or we wish to treat it as if it were (i.e., where data collection rounds are ignored) we can create `Y`, an instance of the `YouthvarsProfile` module, to add minimal metadata (the name of the  unique identifier variable).

```{r}
Y <- YouthvarsProfile(a_Ready4useDyad = X, id_var_nm_1L_chr = "fkClientID")
```

If the temporal dimension of the dataset is important, it may be therefore preferable to instead transform `X` into a `YouthvarsSeries` module instance.  `YouthvarsSeries` objects contain all of the fields of `YouthvarsProfile` objects, but also include additional fields that are specific for longitudinal datasets (e.g. `timepoint_var_nm_1L_chr` and `timepoint_vals_chr` that respectively specify the data-collection timepoint variable name and values and `participation_var_1L_chr` that specifies the desired name of a yet to be created variable that will summarise the data-collection timepoints for which each unit record supplied data).

```{r}
Z <- YouthvarsSeries(a_Ready4useDyad = X,
                     id_var_nm_1L_chr = "fkClientID",
                     participation_var_1L_chr = "participation",
                     timepoint_vals_chr = c("Baseline","Follow-up"),
                     timepoint_var_nm_1L_chr = "round")
```

## YouthvarsProfile methods
### Inspect data
We can now specify the variables that we would like to prepare descriptive statistics for by using the `renew` method. The variables to be profiled are specified in the `profile_chr` argument, the number of decimal digits (default = 3) of numeric values in the summary tables to be generated can be specified with `nbr_of_digits_1L_int`.

```{r}
Y <- renew(Y, nbr_of_digits_1L_int = 2L, profile_chr = c("d_age","d_sexual_ori_s","d_studying_working"))
```

We can now view the descriptive statistics we created in the previous step.

```{r overalltb, tab.cap='Descriptive summary', tab.id = 'overalltb', results="asis"}
Y %>%
  exhibit(profile_idx_int = 1L, scroll_box_args_ls = list(width = "100%"))
```

We can also plot the distributions of selected variables in our dataset.

```{r totalscsnlfig, fig.cap="SOFAS total scores", out.width = "600px"}
depict(Y, x_vars_chr = c("c_sofas"), x_labels_chr = c("SOFAS"), as_percent_1L_lgl = T, style_1L_chr = "lancet" ,what_1L_chr = "histogram", bins = 10)
```


## YouthvarsSeries methods
### Validate data
To explore longitudinal data we need to first use the `ratify` method to ensure that `Z` has been appropriately configured for methods examining datasets reporting measures at two timepoints.

```{r}
Z <- ratify(Z,
            type_1L_chr = "two_timepoints")
```

### Inspect data
We can now specify the variables that we would like to prepare descriptive statistics for using the `renew` method. The variables to be profiled are specified in arguments beginning with "compare_". Use `compare_ptcpn_chr` to compare variables based on whether cases reported data at one or both timepoints and `compare_by_time_chr` to compare the summary statistics of variables by timepoints, e.g at baseline and follow-up. If you wish these comparisons to report p values, then use the `compare_ptcpn_with_test_chr` and `compare_by_time_with_test_chr` arguments.

```{r}
Z <- renew(Z,
           compare_by_time_chr = c("d_age","d_sexual_ori_s","d_studying_working"),
           compare_by_time_with_test_chr = c("k6_total", "phq9_total", "bads_total"),
           compare_ptcpn_with_test_chr = c("k6_total", "phq9_total", "bads_total")) 
```

The tables generated in the preceding step can be inspected using the `exhibit` method.

```{r participtb, tab.cap='Outcomes by data completeness', tab.id = 'participtb', results="asis"}
Z %>%
  exhibit(profile_idx_int = 1L,
          scroll_box_args_ls = list(width = "100%"))
```

```{r outcomestb, tab.cap='Outcomes by data collection round', tab.id = 'outcomestb', results="asis"}
Z %>%
  exhibit(profile_idx_int = 2L,
          scroll_box_args_ls = list(width = "100%"))
```

```{r outcomestbtst, tab.cap='Outcomes by data collection round (with p values)', tab.id = 'outcomestbtst', results="asis"}
Z %>%
  exhibit(profile_idx_int = 3L,
          scroll_box_args_ls = list(width = "100%"))
```

The `depict` method can create plots, comparing numeric variables by timepoint.

```{r totalsfig, fig.cap="SOFAS total scores by data collection round", out.width = "600px"}
depict(Z, x_vars_chr = c("c_sofas"), x_labels_chr = c("SOFAS"), y_labels_chr = "", z_vars_chr = "round",
       z_labels_chr = "Time", as_percent_1L_lgl = T, position_xx  ="dodge", style_1L_chr = "lancet",
       what_1L_chr = "histogram", bins=10)
```

## Share data
**If and only if the dataset you are working with is appropriate for public dissemination (e.g. is synthetic data), you can use the following workflow for sharing it.**
We can share the dataset we created for this example using the `share` method, specifying the repository to which we wish to publish the dataset (and for which we have write permissions) in a ([Ready4useRepos object](https://ready4-dev.github.io/ready4use/articles/V_01.html)). 

```{r eval = F}
A <- Ready4useRepos(gh_repo_1L_chr = "ready4-dev/youthvars", # Replace with your repository 
                          gh_tag_1L_chr = "Documentation_0.0"), # (need write permissions).
A <- share(A,
           obj_to_share_xx = Z,
           fl_nm_1L_chr = "ymh_YouthvarsSeries")
```

`Z` is now available for download as the file `ymh_YouthvarsSeries.RDS` from the ["Documentation_0.0" release of the youthvars package](https://github.com/ready4-dev/youthvars/releases/tag/Documentation_0.0).
