---
title: "Describing Longitudinal Datasets"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Describing Longitudinal Datasets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE}
library(youthvars)
library(ready4use)
```

Youthvars provides a number of tools that are useful for describing features of longitudinal datasets.

## Ingest data
The tools in `youthvars` extend classes and methods from the [ready4use](https://ready4-dev.github.io/ready4use/) package. Our starting point is `x`, an object containint a `Ready4useDyad` (a dataset and data dictionary pair) that we can download from a remote repository.

```{r}
x <- ready4use::Ready4useRepos(dv_nm_1L_chr = "fakes",
                               dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/W95KED",
                               dv_server_1L_chr = "dataverse.harvard.edu") %>%
  ingest(fls_to_ingest_chr = "ymh_clinical_dyad_r4")
```

## Add metadata
We can now add metadata about `x`, transforming the object into an instance of the `YouthvarsSeries` class that is appropriate for longitudinal datasets.
```{r}
x <- YouthvarsSeries(a_Ready4useDyad = x@b_Ready4useIngest@objects_ls$ymh_clinical_dyad_r4,
                     id_var_nm_1L_chr = "fkClientID",
                     participation_var_1L_chr = "participation",
                     timepoint_vals_chr = c("Baseline","Follow-up"),
                     timepoint_var_nm_1L_chr = "round")
```

## Validate data
We next use the `ratify` method to ensure that `x` is appropriately configured for methods examining datasets reporting measures at two timepoints.

```{r}
x <- ratify(x,
            type_1L_chr = "two_timepoints")
```

## Inspect data
We can now specify the variables that we would like to prepare descriptive statistics for using the `renewSlot` and `renew` methods. The variables to be profiled are specified by passing a list to the `profiled_vars_ls` argument, with each element of that list a set of variables for a type of descriptive summary. Currently, the type of descriptive summaries that can be prepared for a `YouthvarsSeries` instance are "participation" (comparison based on whether cases reported data at one or both timepoints) and "temporal" (comparison based on timepoints, e.g at baseline and follow-up). The extension "_tested" is used if p values are to be reported.

```{r}
x <- renewSlot(x,
               profiled_vars_ls = list(temporal = c("d_age","d_sexual_ori_s","d_ATSI","d_studying_working","d_relation_s"),  
                                       temporal_tested = c("k6_total", "phq9_total", "bads_total", "gad7_total"), 
                                       participation_tested = c("k6_total", "phq9_total", "bads_total", "gad7_total")),
               slot_nm_1L_chr = "descriptives_ls")
x <- renew(x,
           type_1L_chr = "characterize")

```

The tables generated in the preceding step can be inspected using the `exhibit` method.

```{r participtb, tab.cap='Outcomes by data completeness', tab.id = 'participtb', results="asis"}
x %>%
  exhibit(type_1L_chr = "characterize",
          profile_idx_int = 1L,
          output_type_1L_chr = "HTML")
```

```{r outcomestb, tab.cap='Outcomes by data collection round', tab.id = 'outcomestb', results="asis"}
x %>%
  exhibit(type_1L_chr = "characterize",
          profile_idx_int = 2L,
          output_type_1L_chr = "HTML")
```

```{r outcomestbtst, tab.cap='Outcomes by data collection round (with p values)', tab.id = 'outcomestbtst', results="asis"}
x %>%
  exhibit(type_1L_chr = "characterize",
          profile_idx_int = 3L,
          output_type_1L_chr = "HTML")
```
The `depict` method can create plots, comparing numeric variables by timepoint.


```{r totalsfig, fig.cap="SOFAS total scores by data collection round", out.width = "600px"}
depict(x,
       type_1L_chr = "by_time",
       var_nms_chr = "c_sofas",
       label_fill_1L_chr = "Time",#
       labels_chr = "SOFAS",#
       y_label_1L_chr = "")
```

