---
title: "Describe And Disseminate Youth Mental Health Datasets"
output: 
  rmarkdown::html_vignette:
    toc: true
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{Describe And Disseminate Youth Mental Health Datasets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

Note: **This vignette is illustrated with fake data**. The dataset explored in this example should not be used to inform decision-making.

```{r echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r results='hide', message=FALSE, warning=FALSE}
library(ready4)
library(youthvars)
```

Youthvars provides a two classes - `YouthvarsProfile` and `YouthvarsSeries` that are useful for describing features of datasets. The tools in `youthvars` build on the metadata included in a [Ready4useDyad](https://ready4-dev.github.io/ready4use/articles/V_02.html).

## Ingest data
To start we ingest  `X`, a `Ready4useDyad` (dataset and data dictionary pair) that we can download from a remote repository.

```{r}
X <- ready4use::Ready4useRepos(dv_nm_1L_chr = "fakes",
                               dv_ds_nm_1L_chr = "https://doi.org/10.7910/DVN/W95KED",
                               dv_server_1L_chr = "dataverse.harvard.edu") %>%
  ingest(fls_to_ingest_chr = "ymh_clinical_dyad_r4",
         metadata_1L_lgl = F)
```

## Add metadata
We could add metadata about `X`, such as the unique identifier variable name, by transforming it to a `YouthvarsProfile` instance.

```{r}
## Not run
# X <- YouthvarsProfile(a_Ready4useDyad = X,
#                       id_var_nm_1L_chr = "fkClientID")
```

However, in this case the data we ingested includes a longitudinal dataset. It is therefore preferable to transform `X` into a `YouthvarsSeries` instance.  `YouthvarsSeries` objects contain all of the fields of `YouthvarsProfile` objects, but also include additional fields that are specific for longitudinal datasets (e.g. `timepoint_var_nm_1L_chr` and `timepoint_vals_chr` that respectively specify the data-collection timepoint variable name and values and `participation_var_1L_chr` that specifies the desired name of a yet to be created variable that will summarise the data-collection timepoints for which each unit record supplied data).

```{r}
X <- YouthvarsSeries(a_Ready4useDyad = X,
                     id_var_nm_1L_chr = "fkClientID",
                     participation_var_1L_chr = "participation",
                     timepoint_vals_chr = c("Baseline","Follow-up"),
                     timepoint_var_nm_1L_chr = "round")
```

## YouthvarsSeries methods
Currently, only methods for `YouthvarsSeries` (and not yet `YouthvarsProfile`) have been included with the `youthvars` package. These methods are summarised in the following sections.

### Validate data
We use the `ratify` method to ensure that `X` has been appropriately configured for methods examining datasets reporting measures at two timepoints.

```{r}
X <- ratify(X,
            type_1L_chr = "two_timepoints")
```

### Inspect data
We can now specify the variables that we would like to prepare descriptive statistics for using the `renewSlot` and `renew` methods. The variables to be profiled are specified in arguments beginning with "compare_". Use `compare_ptcpn_chr` to compare variables based on whether cases reported data at one or both timepoints and `compare_by_time_chr` to compare the summary statistics of variables by timepoints, e.g at baseline and follow-up. If you wish these comparisons to report p values, then use the `compare_ptcpn_with_test_chr` and `compare_by_time_with_test_chr` arguments.

```{r}
X <- renewSlot(X,
               "descriptives_ls",
               compare_by_time_chr = c("d_age","d_sexual_ori_s","d_studying_working"),
               compare_by_time_with_test_chr = c("k6_total", "phq9_total", "bads_total"),
               compare_ptcpn_with_test_chr = c("k6_total", "phq9_total", "bads_total")) %>%
  renew(type_1L_chr = "characterize")

```

The tables generated in the preceding step can be inspected using the `exhibit` method.

```{r participtb, tab.cap='Outcomes by data completeness', tab.id = 'participtb', results="asis"}
X %>%
  exhibit(profile_idx_int = 1L)
```

```{r outcomestb, tab.cap='Outcomes by data collection round', tab.id = 'outcomestb', results="asis"}
X %>%
  exhibit(profile_idx_int = 2L)
```

```{r outcomestbtst, tab.cap='Outcomes by data collection round (with p values)', tab.id = 'outcomestbtst', results="asis"}
X %>%
  exhibit(profile_idx_int = 3L)
```

The `depict` method can create plots, comparing numeric variables by timepoint.

```{r totalsfig, fig.cap="SOFAS total scores by data collection round", out.width = "600px"}
depict(X,
       type_1L_chr = "by_time",
       var_nms_chr = c("c_sofas"),
       label_fill_1L_chr = "Time",#
       labels_chr = c("SOFAS"),#
       y_label_1L_chr = "")
```

## Share data
**If and only if the dataset you are working with is appropriate for public dissemination (e.g. is synthetic data), you can use the following workflow for sharing it.**
We can share the dataset we created for this example using the `share` method, specifying the repository to which we wish to publish the dataset (and for which we have write permissions) in a ([Ready4useRepos object](https://ready4-dev.github.io/ready4use/articles/V_01.html)). 

```{r eval = F}
Y <- Ready4useRepos(gh_repo_1L_chr = "ready4-dev/youthvars", # Replace with your repository 
                          gh_tag_1L_chr = "Documentation_0.0"), # (need write permissions).
Y <- share(Y,
           obj_to_share_xx = X,
           fl_nm_1L_chr = "ymh_YouthvarsSeries")
```

`X` is now available for download as the file `ymh_YouthvarsSeries.RDS` from the ["Documentation_0.0" release of the youthvars package](https://github.com/ready4-dev/youthvars/releases/tag/Documentation_0.0).
